---
title: "ThreeT2DcriteriaAssociationBIMWAIST"
output: html_document
theme: cerulean
---
<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

######in this document, by the same way of thinking to make the T2D case and control in each population ( three T2D criteria, and obviously more individuals as Case),I am calculating association between T2D and BMI, WAIST, WEIGHT and HEIGHT in a logistic regressin adjusted by SEX, AGE and PCA. Then I wanted to see how BMI and WAIST could associate with T2D separately and together. Also I did check about the weight and height association with T2D as well in each populatio. Overal result shows that BMI is not associated with T2D in EastPolynesian(p= 0.653) and Westpolynesian (weack association, p=0.020) wherase WAIST is strongly associated with T2D and this is the core of the finding of this study so far. I will discuss it in depth somewheerelse (result section of the manuscript). by taking out WAIST from the model BMI has the strong associated with T2D. I did some playing around the hight and weight as well and I show the result in the (result section of th emanuscript) 

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(narnia)


AllNewCleanPopulation<-read.delim(  "../Data/AllNewCleanDatasetAndPCA.txt"  , header=TRUE, na.strings=c("NA",""," "))
```
###Eastpolynesian population includind case and control 
```{r }

AllNewCleanPopulation%>% select(PCAETHBROAD,DIABETES)%>% table(., useNA='always')%>% cbind()

EastPolynesiancasecontrol<-subset(AllNewCleanPopulation,Ourcase %in% c("East Polynesian"), value=TRUE)

EastPolynesiancasecontrol%>% select (DIABETES) %>% table(., useNA='always')%>% cbind() 


#outliers for East polynesian continous dataset

vars2<-EastPolynesiancasecontrol[,c("AGECOL","WAIST","HEIGHT","WEIGHT","BMI","SYSTOLIC","AGECOL","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]

vars2nooutlier<-EastPolynesiancasecontrol[,c("PATIENTID","WAIST","WEIGHT","HEIGHT","BMI","SYSTOLIC","AGECOL","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]



j <- 0
for (i in vars2){
  j <- j+1
  mn <- mean(i, na.rm=T)
  ss <- sd(i, na.rm=T)
  outliers <- i > (mn + 4*ss) | i < (mn - 4*ss)
  table(outliers)
  var_range <- c(min(mn - 4*ss, min(i, na.rm=T)),
                 max(mn + 4*ss, max(i, na.rm=T)))
  hist(i, xlim=var_range, main=names(vars2)[j])
  abline(v = (mn + 4*ss))
  abline(v = (mn - 4*ss))
  abline(v =mn, col='blue')
  cat(names(vars2)[j], "\n")
  outliers %>%
    which() %>% 
    print()
  cat('\n')
}



EastPolynesiancasecontrolNoOutlier<-EastPolynesiancasecontrol[-c(1189,153,454,344,797,1188,1241,
                                                                 127,918,1310,1050,309,633,900,941,958,1015,1044,1152,1228,1284,365,894,959,515,88,794),]


EastPolynesiancasecontrolNoOutlier %>% select (DIABETES) %>% table(., useNA='always')%>% cbind() 

###East polynesiancase, Based on theree criteria column (DIABETES, DIABTREAT,DIABINCULDEB) column ( it means any individual with just one of the positive of these three criteria is diabetic case)

Eastpolynesiancase<-EastPolynesiancasecontrolNoOutlier%>% select (PATIENTID,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                  HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                  CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR,PCA1,PCA2,
                                                                  PCA3,PCA4,PCA5,PCA6,PCA7,PCA8,PCA9,PCA10,PCAETHBROAD) %>% filter ((DIABETES==2)|(DIABINCULDEB==2)|(DIABETESTREAT!="T1D treatment"))



#East polynesiancontrol

Eastpolynesiancontrol<-anti_join(EastPolynesiancasecontrolNoOutlier,Eastpolynesiancase)%>% select (PATIENTID,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                  HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                  CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR,PCA1,PCA2,
                                                                  PCA3,PCA4,PCA5,PCA6,PCA7,PCA8,PCA9,PCA10,PCAETHBROAD)
wholeEastpolynesianPopulation<-rbind(Eastpolynesiancontrol,Eastpolynesiancase)

#####association of T2D and other variables

DIABETES<-ifelse(wholeEastpolynesianPopulation$DIABETES=="2",1,0)
wholeEastpolynesianPopulation[,"DIABETES"]<-c(DIABETES)

model<-glm(DIABETES~AGECOL+BMI+WAIST+as.factor(SEX)
            +as.factor(GOUTAFFSTAT)+PCA1+PCA2+
              PCA3+PCA4+PCA5+PCA6+PCA7+PCA8+PCA9+PCA10,
           family='binomial',data=wholeEastpolynesianPopulation)
summary(model)

coef(summary(model))[,'Pr(>|z|)']

###Removing Waist makes BMI significant means WAIST is a strong associated factore to T2D when it is in a model it cancelles BMI effect
model1<-glm(DIABETES~AGECOL+BMI+as.factor(SEX)
            +as.factor(GOUTAFFSTAT)+PCA1+PCA2+
              PCA3+PCA4+PCA5+PCA6+PCA7+PCA8+PCA9+PCA10,
            family='binomial',data=wholeEastpolynesianPopulation)
summary(model1)


###addinh Hight and Weigth

model1HeightWeight<-glm(DIABETES~AGECOL+ HEIGHT+WEIGHT+WAIST+BMI+as.factor(SEX)
            +as.factor(GOUTAFFSTAT)+PCA1+PCA2+
              PCA3+PCA4+PCA5+PCA6+PCA7+PCA8+PCA9+PCA10,
            family='binomial',data=wholeEastpolynesianPopulation)
summary(model1HeightWeight)

coef(summary(model1HeightWeight))[,'Pr(>|z|)']

```
###Westpolynesian population including case and control 
```{r }
WestPolynesiancasecontrol<-subset(AllNewCleanPopulation,Ourcase %in% c("West Polynesian"), value=TRUE)

WestPolynesiancasecontrol %>% select (DIABETES) %>% table(., useNA='always')%>% cbind()

#outliers for West polynesian continoud dataset
vars3<-WestPolynesiancasecontrol[,c("AGECOL","WAIST","HEIGHT","WEIGHT","BMI","SYSTOLIC","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]
vars3nooutlier<-WestPolynesiancasecontrol[,c("PATIENTID","WAIST","HEIGHT","WEIGHT","BMI","SYSTOLIC","AGECOL","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]


j <- 0
for (i in vars3){
  j <- j+1
  mn <- mean(i, na.rm=T)
  ss <- sd(i, na.rm=T)
  outliers <- i > (mn + 4*ss) | i < (mn - 4*ss)
  table(outliers)
  var_range <- c(min(mn - 4*ss, min(i, na.rm=T)),
                 max(mn + 4*ss, max(i, na.rm=T)))
  hist(i, xlim=var_range, main=names(vars3)[j])
  abline(v = (mn + 4*ss))
  abline(v = (mn - 4*ss))
  abline(v =mn, col='blue')
  cat(names(vars3)[j], "\n")
  outliers %>%
    which() %>% 
    print()
  cat('\n')
}


WestPolynesiancasecontrolNoOutlier<-WestPolynesiancasecontrol[-c(392,468,491,128,602,381,476,597,714,715,793,77,378,525,793,796),]


WestPolynesiancasecontrolNoOutlier %>% select (DIABETES) %>% table(., useNA='always')%>% cbind() 




Westpolynesiancase<-WestPolynesiancasecontrolNoOutlier%>% select (PATIENTID,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                  HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                  CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR,PCA1,PCA2,
                                                                     PCA3,PCA4,PCA5,PCA6,PCA7,PCA8,PCA9,PCA10,PCAETHBROAD) %>%  filter  ((DIABETES==2)|(DIABINCULDEB==2)|(DIABETESTREAT!="Untreated"))


#West polynesiancontrol

Westpolynesiancontrol<-anti_join(WestPolynesiancasecontrolNoOutlier,Westpolynesiancase) %>% select (PATIENTID,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                  HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                  CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR,PCA1,PCA2,
                                                                     PCA3,PCA4,PCA5,PCA6,PCA7,PCA8,PCA9,PCA10,PCAETHBROAD)

wholeWestpolynesianPopulation<-rbind(Westpolynesiancontrol,Westpolynesiancase)

###Association T2D and other variables

DIABETES<-ifelse(wholeWestpolynesianPopulation$DIABETES=="2",1,0)
wholeWestpolynesianPopulation[,"DIABETES"]<-c(DIABETES)

model2a<-glm(DIABETES~AGECOL+WAIST+BMI+as.factor(SEX)
            +as.factor(GOUTAFFSTAT)+PCA1+PCA2+
              PCA3+PCA4+PCA5+PCA6+PCA7+PCA8+PCA9+PCA10,
            family='binomial',data=wholeWestpolynesianPopulation)
summary(model2a)
coef(summary(model2a))[,'Pr(>|z|)']

###removing WAIST in this population doesnt  make that much effect on BMI
model2b<-glm(DIABETES~AGECOL+BMI+as.factor(SEX)
            +as.factor(GOUTAFFSTAT)+PCA1+PCA2+
              PCA3+PCA4+PCA5+PCA6+PCA7+PCA8+PCA9+PCA10,
            family='binomial',data=wholeWestpolynesianPopulation)
summary(model2b)


#####adding wight and hight
model2HeightWeight<-glm(DIABETES~AGECOL+ HEIGHT+WEIGHT+WAIST+BMI+as.factor(SEX)
            +as.factor(GOUTAFFSTAT)+PCA1+PCA2+
              PCA3+PCA4+PCA5+PCA6+PCA7+PCA8+PCA9+PCA10,
            family='binomial',data=wholeWestpolynesianPopulation)
summary(model2HeightWeight)

coef(summary(model2HeightWeight))[,'Pr(>|z|)']



```
###European population including case and control
```{r}
Europeancasecontrol<-subset(AllNewCleanPopulation,Ourcase %in% c("European"), value=TRUE)
Europeancasecontrol %>% select (DIABETES) %>% table(., useNA='always')%>% cbind() 

#outliers for European population continous dataset

vars4<-Europeancasecontrol[,c("AGECOL","WAIST","HEIGHT","WEIGHT","BMI","SYSTOLIC","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]

vars4nooutlier<-Europeancasecontrol[,c("PATIENTID","WAIST","HEIGHT","BMI","SYSTOLIC","AGECOL","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]

j <- 0
for (i in vars4){
  j <- j+1
  mn <- mean(i, na.rm=T)
  ss <- sd(i, na.rm=T)
  outliers <- i > (mn + 4*ss) | i < (mn - 4*ss)
  table(outliers)
  var_range <- c(min(mn - 4*ss, min(i, na.rm=T)),
                 max(mn + 4*ss, max(i, na.rm=T)))
  hist(i, xlim=var_range, main=names(vars4)[j])
  abline(v = (mn + 4*ss))
  abline(v = (mn - 4*ss))
  abline(v =mn, col='blue')
  cat(names(vars4)[j], "\n")
  outliers %>%
    which() %>% 
    print()
  cat('\n')
}

#finding PATIENTID to remove outliers


EuropeancasecontrolNoOutlier<-Europeancasecontrol[-c(25,29,174,669,1489,544,757,1429,148,154,170,187,608,1146,1271,1589,1482,1503,1206,561,874,877,946,1024,1139,1164,1188,1256,1368,1395,1480,1531,1585,1659,532,910,1268,1568,1633,490,730,80,1167,1571,1592,1604,1626,1684
                                                     ),]

EuropeancasecontrolNoOutlier %>% select (DIABETES) %>% table(., useNA='always')%>% cbind()

###European Case
Europeancase<-EuropeancasecontrolNoOutlier%>% select (PATIENTID,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                  HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                  CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR,PCA1,PCA2,
                                                      PCA3,PCA4,PCA5,PCA6,PCA7,PCA8,PCA9,PCA10,PCAETHBROAD)  %>% filter ((DIABETES==2)|(DIABINCULDEB==2)|(DIABETESTREAT!="T1D treated with medication"))



#European control

###European control
Europeancontrol<-anti_join(EuropeancasecontrolNoOutlier,Europeancase)  %>% select (PATIENTID,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                  HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                  CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR,PCA1,PCA2,
                                                      PCA3,PCA4,PCA5,PCA6,PCA7,PCA8,PCA9,PCA10,PCAETHBROAD) 

wholeEuropeanPopulation<-rbind(Europeancontrol,Europeancase)

DIABETES<-ifelse(wholeEuropeanPopulation$DIABETES=="2",1,0)

wholeEuropeanPopulation[,"DIABETES"]<-c(DIABETES)

model3<-glm(DIABETES~AGECOL+WAIST+BMI+as.factor(SEX)
            +as.factor(GOUTAFFSTAT)+PCA1+PCA2+
              PCA3+PCA4+PCA5+PCA6+PCA7+PCA8+PCA9+PCA10,
            family='binomial',data=wholeEuropeanPopulation)
summary(model3)

coef(summary(model3))[,'Pr(>|z|)']



```
###Polynesian population including case and control  
```{r }
Polynesiancasecontrol<-subset(AllNewCleanPopulation,Ourcase %in% c("Polynesian"), value=TRUE)

Polynesiancasecontrol %>% select (DIABETES) %>% table(., useNA='always')%>% cbind()

#outliers for  polynesian continoud dataset

vars5<-Polynesiancasecontrol[,c("AGECOL","WAIST","HEIGHT","WEIGHT","BMI","SYSTOLIC","AGECOL","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]
vars5nooutlier<-Polynesiancasecontrol[,c("PATIENTID","WAIST","HEIGHT","WEIGHT","BMI","SYSTOLIC","AGECOL","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]

j <- 0
for (i in vars5){
  j <- j+1
  mn <- mean(i, na.rm=T)
  ss <- sd(i, na.rm=T)
  outliers <- i > (mn + 4*ss) | i < (mn - 4*ss)
  table(outliers)
  var_range <- c(min(mn - 4*ss, min(i, na.rm=T)),
                 max(mn + 4*ss, max(i, na.rm=T)))
  hist(i, xlim=var_range, main=names(vars5)[j])
  abline(v = (mn + 4*ss))
  abline(v = (mn - 4*ss))
  abline(v =mn, col='blue')
  cat(names(vars5)[j], "\n")
  outliers %>%
    which() %>% 
    print()
  cat('\n')
}

#finding PATIENTID to remove outliers

PolynesiancasecontrolNoOutlier<-Polynesiancasecontrol[-c(92,99,85,103,60),]

PolynesiancasecontrolNoOutlier %>% select (DIABETES) %>% table(., useNA='always')%>% cbind()



###Mixed-Polynesian Case
Polynesiancase<-PolynesiancasecontrolNoOutlier%>% select (PATIENTID,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                  HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                  CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR,PCA1,PCA2,
                                                          PCA3,PCA4,PCA5,PCA6,PCA7,PCA8,PCA9,PCA10,PCAETHBROAD) %>% filter ((DIABETES==2)|(DIABINCULDEB==2)|(DIABETESTREAT!="T1D treated with medication"))


#Polynesian control

Polynesiancontrol<-anti_join(PolynesiancasecontrolNoOutlier,Polynesiancase)  %>% select (PATIENTID,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                  HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                  CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR,PCA1,PCA2,
                                                          PCA3,PCA4,PCA5,PCA6,PCA7,PCA8,PCA9,PCA10,PCAETHBROAD)

wholePolynesianPopulation<-rbind(Polynesiancontrol,Polynesiancase)

DIABETES<-ifelse(wholePolynesianPopulation$DIABETES=="2",1,0)
wholePolynesianPopulation[,"DIABETES"]<-c(DIABETES)

model4<-glm(DIABETES~AGECOL+WAIST+BMI+as.factor(SEX)
            +as.factor(GOUTAFFSTAT)+PCA1+PCA2+
            PCA3+PCA4+PCA5+PCA6+PCA7+PCA8+PCA9+PCA10,
            family='binomial',data=wholePolynesianPopulation)
summary(model4)

coef(summary(model4))[,'Pr(>|z|)']

```