---
title: "TrhreeCriteriaT2DMasterDataset"
output: html_document
theme: cerulean
---
<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

######In this version of my analysing I have used the phenotype datasets (4 datasets). This analysisng is one way of thinking to make up the case and control for T2D based on the tree criteria (which have different column names in the 4 database but mainly is self/GP report for having T2D, under treatment/ medicine related to T2D and Randomglocuse.) in this version of analysing, I have made the case based on (or |) comand in filterring for three columns and then checked it by venn diagramm. the T2D case and control for the four datasets after , age<18, NA for T2D, QC filterring QC are as belwo. They have been used to make the master dataset:

###DMgenotypeQCindividualspassedQC,GOUTgenotypeQCindividualspassedQC,RDgenotypeQCindividualspassedQC,NPHgenotypeQCindividualspassedQC.Individuals with DIABETES column= Yes, which means they have one criteria (self/GP)= 180, Individuals with DIABETES column= No, which means they have one criteria (self/GP)= 58 for DMDatsetQCPassed which reads the DMgenotypeQCindividualspassedQC dataset (made after filterring for age<18, NA for T2D and QC) 
```{r }
library(dplyr)
library(gplots)

DMDatasetQCPassed<-read.delim("../Data/DMgenotypeQCindividualspassedQC.txt" , header = T, na.strings = c(""," ","NA"))

DMDatasetQCPassed %>% select (DIABETES) %>% table(., useNA='always') %>% cbind()
```
### Making T2DCase= 258 and T2DControl= 9 for DM dataset
```{r }

#T2DControl

DMControl<-DMDatasetQCPassed %>% select (PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                  HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                  CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR
) %>% filter ((DIABETES==1),(DIABINCULDEB==1), is.na(DIABETESTREAT))


write.table (DMControl,"../Data/DMControl.txt",sep="\t", row.names=F, quote = F)


#T2DCase


DMCase<-anti_join(DMDatasetQCPassed,DMControl)  %>% select (PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                  HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                  CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR
)%>% filter ((DIABETES==2)|(DIABINCULDEB==2)|(DIABETESTREAT!= "T1D Treatment"))

write.table (DMCase,"../Data/DMCase.txt",sep="\t", row.names=F, quote = F)


T2DDM<-DMCase

T2DDM %>% select (DIABETES) %>% table(., useNA='always') %>% cbind()

x<-DMDatasetQCPassed %>% select (PATIENT,DIABETES)%>% filter (DIABETES==2)
x<-x$PATIENT

y<-DMDatasetQCPassed %>% select (PATIENT,DIABINCULDEB)%>% filter (DIABINCULDEB==2)
y<-y$PATIENT

z<-DMDatasetQCPassed %>% select (PATIENT,DIABETESTREAT)%>% filter (DIABETESTREAT!= "T1D Treatment")
z<-z$PATIENT
samples<-list( DIABETESyes=x,DIABINCLUDEByes=y, TREATyes=z)
venn(samples)
```
### borderline=3 means maybe T2D =4. Individuals with DIABETES column= Yes, which means they have one criteria (self/GP)= 584, Individuals with DIABETES column= No, which means they have one criteria (self/GP)= 2970 for GoutDatasetQCPassed which reads the GOUTgenotypeQCindividualspassedQC dataset (made after filterring for age<18, NA for T2D and QC
```{r }
GoutDatasetQCPassed<-read.delim("../Data/GOUTgenotypeQCindividualspassedQC.txt" , header = T, na.strings = c(""," ","NA"))
GoutDatasetQCPassed %>% select (DIABETES) %>% table()

```
### Making T2DCase=1272  and Control= 2292  separately for GOUT dataset
```{r }
GOUTData<-read.delim("../Data/GOUTgenotypeQCindividualspassedQC.txt"  , header=TRUE, na.strings=c("NA",""," "))

GOUTCase<-GOUTData %>% select(PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                      HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GPGOUTAFFSTAT,GLUCMEASURED,
                                      CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR_SCL
) %>%  filter ((DIABETES==2)|(GLUCMEASURED==2)|(DIABETESTREAT!= "T1D Treatment"))

###GOUTDataset case with Venn diagram for three criteria

x<-GOUTData %>% select (PATIENT,DIABETES)%>% filter (DIABETES==2)
x<-x$PATIENT

y<-GOUTData %>% select (PATIENT,GLUCMEASURED)%>% filter (GLUCMEASURED==2)
y<-y$PATIENT

z<-GOUTData %>% select (PATIENT,DIABETESTREAT)%>% filter (DIABETESTREAT!= "T1D Treatment")
z<-z$PATIENT
samples<-list( DIABETESyes=x,DIABINCLUDEByes=y, TREATyes=z)
venn(samples)


write.table (GOUTCase,"../Data/GOUTCase.txt",sep="\t", row.names=F, quote = F)

#T2DControl

GOUTControl<-anti_join(GOUTData,GOUTCase)%>%select(PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                            HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GPGOUTAFFSTAT,GLUCMEASURED,
                                      CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR_SCL
)

write.table (GOUTControl,"../Data/GOUTControl.txt",sep="\t", row.names=F, quote = F)

T2DGOUT<-rbind(GOUTCase, GOUTControl)

```
###Individuals with DIABETES column= Yes, which means they have one criteria (self/GP)= 204, Individuals with DIABETES column= No, which means they have one criteria (self/GP)= 146 for for RrnalDatsetQCPassed which reads the RDenotypeQCindividualspassedQC dataset (made after filterring for age<18, NA for T2D and QC
```{r }
RenalDatasetQCPassed<-read.delim("../Data/RDgenotypeQCindividualspassedQC.txt" , header = T, na.strings = c(""," ","NA"))
RenalDatasetQCPassed %>% select (DIABETES) %>% table()

```
### RDCase for T2D=208, and RDControl=146
```{r }
#T2DCase

RenalData<-read.delim("../Data/RDgenotypeQCindividualspassedQC.txt"  , header=TRUE, na.strings=c("NA",""," "))


RenalDataCase<-RenalData %>% select(PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,INSULIN,
                                                     HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,HBA1C ,
                                                     CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR
) %>% filter ((DIABETES==2)|(HBA1C>=50)|(INSULIN==1))


write.table (RenalDataCase,"../Data/RenalDataCase.txt",sep="\t", row.names=F, quote = F)
#T2DControl

RenalDataControl<-anti_join(RenalData,RenalDataCase) %>% select(PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,INSULIN,
                                                     HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,HBA1C ,
                                                     CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR
)

write.table (RenalDataControl,"../Data/RenalDataControl.txt",sep="\t", row.names=F, quote = F)

T2DRD<-rbind(RenalDataCase, RenalDataControl)


###RDDataset case with Venn diagram for three criteria

x<-RenalData %>% select (PATIENT,DIABETES)%>% filter (DIABETES==2)
x<-x$PATIENT

y<-RenalData %>% select (PATIENT,HBA1C)%>% filter (HBA1C>=50)
y<-y$PATIENT

z<-RenalData %>% select (PATIENT,INSULIN)%>% filter (INSULIN==1)
z<-z$PATIENT
samples<-list( DIABETESyes=x,DIABINCLUDEByes=y, TREATyes=z)
venn(samples)
```
### Individuals with DIABETES column= Yes, which means they have one criteria (self/GP)= 234, Individuals with DIABETES column= No, which means they have one criteria (self/GP)= 72 for NPHgenotypeQCindividualspassedQC dataset (made after filterring for age<18, NA for T2D and QC)
```{r }
NPHDatasetQCPassed<-read.delim("../Data/NPHgenotypeQCindividualspassedQC.txt" , header = T, na.strings = c(""," ","NA"))
NPHDatasetQCPassed %>% select (DIABETES) %>% table()
```
### NPHT2DCase=77 and Control=230 separately for NPH dataset
```{r }
#T2DCase

NPHData<-read.delim( "../Data/NPHgenotypeQCindividualspassedQC.txt" , header=TRUE, na.strings=c("NA",""," "))


NPHDataCase<-NPHData%>% select (PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR
) %>% filter((DIABETES==2)|(DIABINCULDEB==2)|(DIABETESTREAT!= "T1D Treatment"))

write.table (NPHDataCase,"../Data/NPHDataCase.txt",sep="\t", row.names=F, quote = F)

NPHDataCase %>% select (DIABETES) %>% table(., useNA='always') %>% cbind()

#T2DControl

NPHDataControl<-anti_join(NPHData,NPHDataCase)%>% select(PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR)

write.table (NPHDataControl,"../Data/NPHDataControl.txt",sep="\t", row.names=F, quote = F)

T2DNPH<-rbind(NPHDataCase, NPHDataControl)

###nphDataset case with Venn diagram for three criteria

x<-NPHData%>% select (PATIENT,DIABETES)%>% filter (DIABETES==2)
x<-x$PATIENT

y<-NPHData %>% select (PATIENT,DIABINCULDEB)%>% filter (DIABINCULDEB==2)
y<-y$PATIENT

z<-NPHData %>% select (PATIENT,DIABETESTREAT)%>% filter (DIABETESTREAT!= "T1D Treatment")
z<-z$PATIENT
samples<-list( DIABETESyes=x,DIABINCLUDEByes=y, TREATyes=z)
venn(samples)
```
### Renaming the columns to be the same for rbinding
```{r }
##### making a masterdataset

colnames(T2DDM)

colnames(T2DGOUT)

colnames(T2DRD)

colnames(T2DNPH)



colnames(T2DGOUT)[16] <- "DIABINCULDEB"

colnames(T2DGOUT)[23] <- "EGFR"

colnames(T2DGOUT)[15]<-"GOUTAFFSTAT"


colnames(T2DNPH)[15]<-"GOUTAFFSTAT"

colnames(T2DRD)[7]<-"DIABETESTREAT" 
colnames(T2DRD)[16]<-"DIABINCULDEB"
colnames(T2DRD)[15]<-"GOUTAFFSTAT"


```
### rbinding the datasets and making the masterdatset
```{r }
RenalGOUTDMDNPHataset<-rbind(T2DDM,T2DGOUT,T2DRD,T2DNPH)

write.table (RenalGOUTDMDNPHataset,"../Data/AllNewCleanDatawithduplications.txt",sep="\t", row.names=F, quote = F)

RenalGOUTDMDNPHataset%>% select(DIABETES) %>% table(., useNA='always') %>% cbind()
```
### removing duplicated IDS(repeatedIDs), the output of this rbinding and removing duplications called "AllNewCleanData.txt". 4448 = Datasetbefore  filterring for duplications, 4309= dataset after filterring for duplications.
```{r }
RenalGOUTDMDNPHatasetOrdered<-RenalGOUTDMDNPHataset[order(RenalGOUTDMDNPHataset$DIABETES),]

RenalGOUTDMDNPHatasetNoDuplication<-RenalGOUTDMDNPHatasetOrdered[!duplicated(RenalGOUTDMDNPHatasetOrdered$PATIENT),]

#d<-distinct(RenalGOUTDMDNPHataset)
#d %>%select(DIABETES) %>% table(., useNA='always') %>% cbind()

#b<-unique(RenalGOUTDMDNPHataset)
#b%>%select(DIABETES) %>% table(., useNA='always') %>% cbind()

RenalGOUTDMDNPHatasetNoDuplication%>% select(DIABETES) %>% table(., useNA='always') %>% cbind()

MasterT2DCase<-RenalGOUTDMDNPHatasetNoDuplication %>% filter ((DIABETES==2)|(DIABINCULDEB==2)|(DIABETESTREAT!= "T1D Treatment"))
dim(MasterT2DCase)

MasterT2DControl<-anti_join(RenalGOUTDMDNPHatasetNoDuplication,MasterT2DCase)
dim(MasterT2DControl)
```
### RenalGOUTDMDNPHatasetNoDuplication exporting to Desktop and used as the "AllNewCleanData.txt" file for detecting ancestry populations with PCA.
```{r }
write.table (RenalGOUTDMDNPHatasetNoDuplication,"../Data/AllNewCleanData.txt",sep="\t", row.names=F, quote = F)
```
